---
title: 'Study Analysis'
author: "Markus Hüttner"
output:
  html_document:
    df_print: paged
    theme: readable
    highlight: zenburn
    self_contained: true
---

<style>
.main-container {
  max-width: 95% !important;
  width: 95% !important;
}
</style>

## Libraries

```{r, fig.height=8, fig.width=15, warning=FALSE, message=FALSE}
library(jsonlite)
library(dplyr)
library(tidyr)
library(ggplot2)
library(effsize)
library(rstatix)
```

## Exploratory Data Analysis

### Load and combine data

```{r, fig.height=8, fig.width=15}
base_dir <- getwd()

# Function to load all JSON files from a folder
load_json_folder <- function(folder_path, label) {
  files <- list.files(folder_path, pattern = "\\.json$", full.names = TRUE)
  
  data_list <- lapply(files, function(f) {
    df <- as.data.frame(fromJSON(f))
    df$file <- basename(f)
    return(df)
  })
  
  data <- bind_rows(data_list)
  data$group <- label
  return(data)
}

# Load both groups
data_not_sent <- load_json_folder(file.path(base_dir, "notSentiment"), "NotSentiment")
data_sent <- load_json_folder(file.path(base_dir, "Sentiment"), "Sentiment")

# Combine into one dataset
data_all <- bind_rows(data_not_sent, data_sent)

# Inspect
str(data_all)
print(data_all)
```

### Check for NAs

```{r, fig.height=8, fig.width=15}
colSums(is.na(data_all))
```

### Check Summary

```{r, fig.height=8, fig.width=15}
summary(data_all)
```

### Convert to long format for visualization

```{r, fig.height=8, fig.width=15}
data_long <- data_all %>%
  pivot_longer(cols = -c(file, group), names_to = "question", values_to = "rating")
head(data_long)
```

### Distribution plots

```{r, fig.height=8, fig.width=15}
label_map <- c(
  sat_problem = "Handled problem satisfactorily",
  sat_helpful = "Responses were helpful",
  sat_steps = "Suggested steps were meaningful",
  sat_serious = "Concern was taken seriously",
  sat_overall = "Overall satisfaction",
  
  emp_empathy = "Appeared empathetic",
  emp_mood = "Responded to user's mood",
  emp_tone = "Tone was appropriate",
  emp_underst = "Felt understood",
  emp_support = "Emotionally supportive",
  
  trust_again = "Would trust again",
  reuse = "Would use again",
  prof = "Appeared professional",
  clarity = "Communication was clear"
)

data_long$rating <- factor(data_long$rating, levels = 1:5)
data_long$question <- factor(
  data_long$question,
  levels = names(label_map)
)
data_long$dimension <- dplyr::case_when(
  data_long$question %in% names(label_map)[1:5]  ~ "Task Satisfaction",
  data_long$question %in% names(label_map)[6:10] ~ "Perceived Empathy",
  TRUE ~ "Trust & Overall Impression"
)

p <- ggplot(data_long, aes(x = rating, fill = group)) +
  geom_bar(
    position = position_dodge(width = 0.55, preserve = "single"),
    width = 0.45,
    color = NA,
    alpha = 0.9
  ) +
  facet_wrap(~question,
           ncol = 3,
           labeller = labeller(question = label_map)) +
  scale_fill_manual(
    values = c("#0072B2", "#D55E00"),
    labels = c("Not Sentiment-Aware", "Sentiment-Aware")
  ) +
  scale_x_discrete(drop = FALSE) +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 6)) +
  labs(
    title = "Comparison of Chatbot Ratings",
    x = "Rating",
    y = "Number of Responses",
    fill = "Group"
  ) +
  theme_minimal(base_size = 20) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_line(color = "grey85"),
    
    strip.text = element_text(size = 14),
    
    axis.title.x = element_text(size = 15, margin = margin(t = 8)),
    axis.title.y = element_text(size = 15, margin = margin(r = 8)),
    
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    
    plot.title = element_text(size = 17, face = "bold", hjust = 0.5, margin = margin(b = 10)),
    
    legend.position = "bottom",
    legend.title = element_text(face = "bold"),
    legend.text = element_text(size = 12),
    
    panel.spacing = unit(1.2, "lines")
  )

ggsave("distribution_barplot.png", plot = p,
       width = 10, height = 14, dpi = 300, bg = "white")
```

### Boxplots

```{r, fig.height=8, fig.width=15}
data_long$rating_num <- as.numeric(as.character(data_long$rating))

p <- ggplot(data_long, aes(x = group, y = rating_num, fill = group)) +
  geom_boxplot(
    width = 0.6,
    color = "grey20",
    alpha = 0.9
  ) +
  facet_wrap(~question, ncol = 3,
             labeller = labeller(question = label_map)) +
  scale_fill_manual(
    values = c("#0072B2", "#D55E00"),
    labels = c("Not Sentiment-Aware", "Sentiment-Aware")
  ) +
  labs(
    title = "Comparison of Chatbot Ratings",
    x = "Group",
    y = "Rating",
    fill = "Group"
  ) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_line(color = "grey85"),
    
    strip.text = element_text(size = 14),
    strip.background = element_blank(),
    panel.background = element_blank(),
    
    axis.title.x = element_text(size = 15, margin = margin(t = 8)),
    axis.title.y = element_text(size = 15, margin = margin(r = 8)),
    
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    
    plot.title = element_text(size = 17, face = "bold", hjust = 0.5, margin = margin(b = 10)),
    
    legend.position = "bottom",
    legend.title = element_text(face = "bold"),
    legend.text = element_text(size = 12),
    
    panel.spacing = unit(1.2, "lines")
  )

ggsave("boxplots.png", plot = p,
       width = 10, height = 14, dpi = 300, bg = "white")
```

## Statistical Analysis

### Create Groups

```{r, fig.height=8, fig.width=15}
data_all <- data_all %>%
  mutate(
    satisfaction = rowMeans(select(., starts_with("sat_")), na.rm = TRUE),
    empathy = rowMeans(select(., starts_with("emp_")), na.rm = TRUE),
    trust = rowMeans(select(., trust_again, reuse, prof, clarity), na.rm = TRUE)
  )
```

### Normality tests (Shapiro-Wilk)

```{r, fig.height=8, fig.width=15}
shapiro.test(data_all$satisfaction[data_all$group == "Sentiment"])
shapiro.test(data_all$satisfaction[data_all$group == "NotSentiment"])

shapiro.test(data_all$empathy[data_all$group == "Sentiment"])
shapiro.test(data_all$empathy[data_all$group == "NotSentiment"])

shapiro.test(data_all$trust[data_all$group == "Sentiment"])
shapiro.test(data_all$trust[data_all$group == "NotSentiment"])
```

H0 (null hypothesis): Data is normally distributed
If p >  0.05 -> cannot reject normality
If p <= 0.05 -> data is not normal

### If data is normal then t-test for group comparison

```{r, fig.height=8, fig.width=15}
t.test(satisfaction ~ group, data = data_all)
t.test(empathy ~ group, data = data_all)
# t.test(trust ~ group, data = data_all)
```

### If data is not normal then wilcoxon test for group comparison

```{r, fig.height=8, fig.width=15}
# wilcox.test(satisfaction ~ group, data = data_all)
# wilcox.test(empathy ~ group, data = data_all)
wilcox.test(trust ~ group, data = data_all)
```

H0 (null hypothesis): There is no difference between the groups
If p >  0.05 -> cannot reject H0 (no statistically significant difference)
If p <= 0.05 -> reject H0 (statistically significant difference between groups)

### How big is the difference?

```{r, fig.height=8, fig.width=15}
cohen.d(satisfaction ~ group, data = data_all)
cohen.d(empathy ~ group, data = data_all)
```

```{r, fig.height=8, fig.width=15}
wilcox_effsize(data_all, trust ~ group)
```


